Proof No Diamond (3 layers)

### 2023-10-09 Yuke No Diamond for 3 layers.tex
proof part 1

\begin{theorem}
	If for all $(a,d)$ there is no compensation paths $(\pi,\pi')$ 
	in the network from $a$ to $d$, then the LP approximation is 
	$100\%$ accurate (or deepoly with the $f(x) \geq 0$ abstraction only, never using the 
	$f(x) \geq x$ abstraction). 
\end{theorem}

In this section, we only consider the case of 3 layers: that is, one input layer, one hidden layer and one output layer, which contains only one node. Of course, we assume all active functions are ReLU function. Before we introduce the proof, we list some definitions.

\begin{definition}
	1. We use $a_i$ to denote nodes in the input layer, use $b_j$ to denote nodes in the hidden layer and use $z$ to denote the output node. We use $W$ to denote the weights and $w({ba})$ and $w({zb})$ to denote the components: weight from one node to another. We use capital letter $B$ to denote bias, although it is not important here.

	2. We use $b$ to denote the node before ReLU and $\hat{b}$ to denote the node after ReLU: $\hat{b} = \ReLU(b)$.

@@

\begin{proof}

The following definition is the most important concept in this proof.

\begin{definition}[Sign of Node]
	We define a sign function $S$ on every node $n$ such that: 	(1). $S(n)=0$ if all path from $n$ to $z$ has 0 weight; (2). $S(n)=1$ if all path from $n$ to $z$ has non-negative weight, and at least one path has a positive weight; (3). $S(n)=1$ if all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight.

	More specifically, for an input node $a_i$:

	(1) $S(a_i)=0$ if \begin{align}
		\forall b_j\ w(b_ja_i)\cdot w(zb_j) = 0.
	\end{align}

	(2) $S(a_i)=1$ if\begin{align}
		\forall b_j\ w(b_ja_i)\cdot w(zb_j) \geq 0 \wedge \exists b_j\ w(b_ja_i)\cdot w(zb_j) > 0. 
	\end{align}

	(3) $S(a_i)=1$ if \begin{align}
	\forall b_j\ w(b_ja_i)\cdot w(zb_j) \leq 0 \wedge \exists b_j\ w(b_ja_i)\cdot w(zb_j) < 0. 
\end{align}

And for a node $b_j$ in the hidden layer, $S(b_j)=1,-1,0$ if $w(zb_j)$ is positive, negative, or 0. Finally, for the output node $z$, we define $S(z)=1$.
\end{definition}

Notice that, if there is no Diamond, then $S$ is defined on all nodes, i.e., any input node must be one of above cases (1),(2),(3).

\begin{definition}
	For node $n$, we use $n^s$ to denote $S(n)\cdot n$. And we define that $n^s$ gets maximal value means that: if $S(n)=1$, then $n$ gets maximal value; if $S(n)=-1$, then $n$ gets minimal value (if $S(n)=0$, we say $n^s$ always gets maximal value).

	Similarly we define the meaning of $n^s$ gets minimal value.  
\end{definition}

\begin{lemma}[Sign and Weight]
	1. When one node $b$ (before $\ReLU$ function) gets its maximal or minimal value, $\hat{b}=\ReLU(b)$ also gets maximal or minimal.

	2. Suppose $L$ is a layer and $L'$ is the next layer. Then for any $n$ on $L$ and $m$ on $L'$, if both $S(m)$ and $w(mn)$ are nonzero, then we have that: \begin{align}
		S(n)=S(m)\mathrm{sign}(w(mn))
	\end{align}
\end{lemma}
\begin{proof}
	1. This is obvious.

	2. This can be proved by using definition. If $S(m)$ are nonzero, then there is a path $P$ from $m$ to the output node  $z$ with a nonzero weight of the same sign as $S(m)$.

	Then from $n$ to $m$ and along $P$ to $z$ is a path from $n$ to $z$, and the weight of this path is the product of weight of $P$ and $w(mn)$, which is nonzero and have the same sign as $S(m)\cdot w(mn)$. This is what we want to show.
\end{proof}

The following lemma shows the reason why this theorem is true.

\begin{lemma}
	Suppose $L$ is a layer (before $\ReLU$ function) other than output layer. Then when all $n^s$ gets their maximal value (where $n$ are node on $L$), for the next layer $L'$, and every node $m$ on $L'$, $m^s$ gets its maximal value. In formula: \begin{align}
		\max(m^s)=\sum_{n}w(mn)\ReLU(\max(n^s))+b
	\end{align}

	And for $m^s$ minimal value, we have the same result: \begin{align}
		\min(m^s)=\sum_{n}w(mn)\ReLU(\min(n^s))+b
	\end{align}
\end{lemma}

\begin{proof}
	Assume all $n^s$ on $L$ get their maximal, and $m$ is a node on $L'$, we try to show that $m^s$ gets its maximal value.

	If $S(m)=0$, then we have nothing to do. Without loss of generality,  we assume $S(m)=1$; for $S(m)=-1$, it is highly similar. Then by Lemma 1, for any $n$ in layer $L$ such that $w(mn)\neq 0$, $$S(n)=S(m)\mathrm{sign}(w(mn))=\mathrm{sign}(w(mn)).$$

	On the other hand, $$m=\sum_{n}w(mn)\hat{n}+b.$$ This means if all $\hat{n}$ with $w(mn)>0$ get maximal value and all $\hat{n}$ with $w(mn)<0$ get minimal value, $m$ also gets its maximal value. By Lemma 1, $\hat{n}$ gets maximal value or minimal value if $n$ does. So, when all $n^s$ get maximal value, $m$ will also get its maximal value.

	So, $m^s$ gets maximal if all $n^s$ in previous layer gets their maximal. The part for $m^s$ minimal value is almost the same, by replacing all maximal to minimal.
\end{proof}

From Lemma 2, we can easily see that: when all nodes $a_i^s$ in the input layer get their maximal/minimal value, all nodes $b_j^s$ in the hidden layer get their maximal/minimal value, and then the output node $z$.

The rest part is to show, interval arithmetic, and hence any approximation as accuracy as it, can get these exact bounds.

@@

\end{proof}

>>>>

We assume all active functions are ReLU function.

\begin{definition}

	\begin{itemize}
	 \item  We use $a_i$ to denote nodes in the input layer, use $b_j$ to denote nodes in the first hidden layer, etc and finally use $z_i$ to denote the output nodes.
	For $L,L'$ consecutive layers, for $x \in L$ and $y \in L'$, we denote 
	$w({x y})$ to denote the fixed weight from $x$ to $y$, while we use
	$weight_Y(x)$ to denote the weights of node $x$ from initial configuration $Y$. 
	We use $bias$ to denote bias weightt.

	\item We use $b^-$ to denote the node before ReLU and $b^+$ to denote the node after ReLU: $b^+ = \ReLU(b^-)$.
	\end{itemize}

@@

\begin{theorem}
	\label{th1}
	If for all $(x,y)$ there is no compensation paths $(\pi,\pi')$ 
	in the network from $x$ to $y$, then the LP approximation is 
	$100\%$ accurate (or deepoly with the abstraction $\underline{f}(x) = 0$ for the lower bound, never using the $\underline{f}(x) = x$ abstraction). 
\end{theorem}

\subsection*{Proof of Theorem \ref{th1}}

First, notice that if $weight_Y(b^-)$ is maximal (resp. minimal), 
then $weight_Y(b^+=\ReLU(b^-))$ also gets maximal (resp. minimal).
The following definition is the most important concept in this proof.

\begin{definition}[Sign of Node]
	We define a sign function $S$ on nodes $n$ such that: 	(1). $S(n)=0$ if all path from $n$ to $c$ has 0 weight; (2). $S(n)=1$ if all path from $n$ to $c$ has non-negative weight, and at least one path has a positive weight; (3). $S(n)=-1$ if all path from $n$ to $c$ has non-positive weight, and at least one path has a negative weight. 
	In general, $S$ may not be defined on every node (e.g. if there is a negative and a positive path from $a$ to $c$). However, if there is no Diamond, $S$ is defined on all nodes: any node fulfills one of above cases (1),(2),(3).

	For instance, for an input node $a_i$ of any 3 layer DNN:
	\begin{enumerate}
      \item  $S(a_i)=0$ if 
	  for all $b_j, w(a_i b_j)\cdot w(b_j c) = 0$

	 \item  $S(a_i)=1$ if for all $b_j, w(a_i b_j)\cdot w(b_j c) \geq 0$ and there exists 
	 $j, w(a_i b_j)\cdot w(b_j c) > 0$

	\item $S(a_i)=-1$ if for all $b_j\ w(a_i b_j)\cdot w(b_j c) \leq 0$ and there exists 
	$j, w(a_i b_j)\cdot w(b_j c) < 0$ 
\end{enumerate}

For $b_j$ in the hidden layer, we have $S(b_j)=1,-1,0$ if $w(b_j c)$ is positive, negative, or 0 respectively. Finally, for the output node $c$, we define $S(c)=1$.
\end{definition}

\begin{lemma}[Sign]
	\label{lemma1}
Let $L,L'$  are consecutive layers be of a DNN without compensation. 
Then if both 
$w(m n) \neq 0$ and $S(n) \neq 0$, then 
	$S(m)=S(n)\mathrm{Sign}(w(n m))$.
\end{lemma}

\begin{proof}
	If $S(n) \neq 0$, then there is a path $\pi$ from $n$ to the output node $c$ with a nonzero weight of the same sign as $S(n)$.

	Hence there is a non zero path from $m$ to $c$: $(m n) \pi$, which is of sign 
	$S(n)\mathrm{Sign}(w(mn))$. As there is no compensation, $S(m)=S(n)\mathrm{Sign}(w(mn))$.
\end{proof}

For a node $n$, we use $n_s$ to denote $S(n)\cdot n$. 
Notice that for $S(n)=1$, $n_s$ gets maximal value whenever $n$ gets maximal value; 
while for $S(n)=-1$, $n_s$ gets maximal value whenever $n$ gets minimal value (and vice versa). For $S(n)=0$, $n_s=0$ and thus always reach his minimal and maximal.

\begin{lemma}
	\label{lemma2}
	Let $L,L'$ be consecutive layers of a compensation free DNN, and $n \in L'$. Then:
		$$ \max(n^-_s)=\sum_{m \in L}w(m n) \max(m^+_s)+bias_n \text{ and }$$
		$$\max(n^+_s)=ReLU(\sum_{m \in L}w(m n) max(m^+_s)+bias_n)$$

	Similarly  for minimal value,	
	$ \min(n^-_s)=\sum_{m \in L}w(m n)\min(m^+_s)+bias_n $ and
	$ \min(n^+_s)=ReLU(\sum_{m \in L}w(m n)\min(m^+_s)+bias_n)$
\end{lemma}

\begin{proof}
We choose $\bar{Y}$ maximizing each $a_s \in \{a,-a\}$ in the input layer $L=L_0$,
	that is setting $a$ as $max(a)$ for $S(a)=1$ and $min(a)$ for $S(a)=-1$.
At configuration $\bar{Y}$, every $a_s(\bar{Y}))=\max_Y {a_s(Y))}=max(a_s)$ is maximized.

Consider any $b$ in the next layer $L_1$.
Assume first that $S(b)=+1$.
We have $weight(b_s)= weight(b) = \sum_{a \in L_0} w(a b) weight(a) + bias_b$.
If every $w(a b) weight(a)$ is maximized in the same time, 
then $weight(b_s)= weight(b)$ is also maximized.

We claim that this is the case at 
configuration $\bar{Y}$.
Indeed, if $w(a b)=0$, any $weight(a)$ can be chosen to maximize 
$w(a b) weight(a)$. If $w(a b)>0$ then $S(a) = +1$ by Lemma \ref{lemma1}, since
$S(b)=+1$. We have $max(a_s) = max(a)$, maximizing $w(a b) weight(a)$.
Finally, if $w(a b)<0$ then $S(a_s) = -1$, and we have 
$max(a_s) = -min(a)$, maximizing $w(a b) weight(a)$.

Hence $b$ reaches is maximum in configuration $\bar{Y}$, and the value satisfies: 
$$\max(b)=\sum_{m \in L}w(a b) \max(a_s)+bias_b$$

The case $S(b)=-1$ is symetric:
$\max(b_s)= -\min(b)$ is reached when minimizing $b$, which is also 
satisfied at $\bar{Y}$.
In both case, $weight(b_s)$ is maximized at $\bar{Y}$. Notice that $\bar{Y}$ does not depend upon $b$, so choosing $\bar{Y}$ uniformly maximize all $weight(b_s)$.

This implies that $ReLU(b^-_s)$ is maximized at $\bar{Y}$, and its value is 
$ReLU(\sum_{m \in L}w(a b) \max(a_s)+bias_b)$, again uniformly reached over all $b$ at configuration $\bar{Y}$.

We proceed by induction over every layer of the DNN till the output layer, proving that 
$max(z_s)=max(z)$ is reached at configuration $\bar{Y}$.

The case for $min(x_s)$ is similar, reached at configuration $\underline{Y}$.
\end{proof}

Notice that for all $x$, either $S(x)=+1$ and 
$(\min(x),\max(x))=(\min(x_s),\max(x_s))$, 
or $S(x)=-1$ and $(\min(x),\max(x))=(-\max(x_s),-\min(x_s))$.
In both case, these two bounds are found in Lemma \ref{lemma2},
from configuration $\bar{Y}$ and $\underline{Y}$.
We now show that DeepPoly will generate the same bounds 
$\underline{f}(x)=\min(x)$ and $\bar{f}(x)=\max(x)$.

\paragraph{DeepPoly bounds}

@@
